GMSE: a general tool for management strategy evaluation

### 2017-09-22 notebook/presentation/GMSE_presentation.Rmd
Some additional presentation stuff

# Adaptive management and conservation conflicts

## Process of adaptive management

- Uses an iterative process of decision-making
- Sets explicit management goals
- Develops alternative management strategies
- Deals with uncertainty by updating knowledge of the system

![](African_Bush_Elephant.jpg)
<!--- ![](adaptive_management.png) --->

**Image:** [M. M. Karim](https://en.wikipedia.org/wiki/User:Muhammad_Mahdi_Karim)

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; [Keith et al. (2011) *Biol. Conserv.* **144**: 1175-1778](https://doi.org/10.1016/j.biocon.2010.11.022)

@@

@@

**Conservation conflicts are complex, and characterised by incomplete information that is rapidly changing.**

@@

**Conservation conflicts are complex, and characterised by incomplete information that is rapidly changing.**

@@

- One manager analyses observations and compares with a target abundance
- Five possible policy options (scaring, culling, castration, feeding, helping)
- Fixed budget and a costs for setting policy
- Genetic algorithm anticipates policy effects from previous actions to set costs of actions for stake-holders

@@

# Rise of the geese: Example use of GMSE

- Wild geese declined starting in the 1930s, likely caused by habitat loss and increased hunting
- Refuges were established and hunting restricted in the 1950s, leading to recovery

![](barnacle_goose.JPG)

**Image:** [Raju Kasambe](https://commons.wikimedia.org/wiki/File:Barnacle_Geese_Branta_leucopsis_London_UK_(4).JPG)

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; [Fox & Madsen 2017 *Ambio* **46(s2)**:170-187](http://doi.org/10.1007/s13280-016-0878-2)  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; [Fox & Abraham 2017 *Ambio* **46(s2)**:188-197](http://doi.org/10.1007/s13280-016-0879-1)

- Wild geese declined starting in the 1930s, likely caused by habitat loss and increased hunting
- Refuges were established and hunting restricted in the 1950s, leading to recovery
- Since the 1960s, many populations have exponentially increased
- Population growth coincides with a switch from natural wetlands to agricultural landscapes
- Switch to agricultural exploitation causes conflict with farmers

![](goose_graph.png)

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; [Fox & Madsen 2017 *Ambio* **46(s2)**:170-187](http://doi.org/10.1007/s13280-016-0878-2)  
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; [Fox & Abraham 2017 *Ambio* **46(s2)**:188-197](http://doi.org/10.1007/s13280-016-0879-1)

@@

>>>>

# Uncertainty and conservation conflict

@@

**Conservation conflicts between biodiversity and food security result when the outcome of one stake-holder is affected by one or more other stake-holders** ([Colyvan et al. (2011) *Biol. Conserv.* **144**:1246-1253](http://www.sciencedirect.com/science/article/pii/S000632071000488X?via%3Dihub))

@@

</div>
<br><br>
<center>

**Conservation conflicts between biodiversity and food security result when the outcome of one stake-holder is affected by one or more other stake-holders** ([Colyvan et al. (2011) *Biol. Conserv.* **144**:1246-1253](http://www.sciencedirect.com/science/article/pii/S000632071000488X?via%3Dihub))

@@

@@

- One manager analyses observations and compares with a target abundance
- Five possible policy options (scaring, culling, castration, feeding, helping)
- Fixed budget and costs for setting policy
- Genetic algorithm anticipates policy effects from previous actions to set costs of actions for stake-holders

@@

# Example use of GMSE

@@

## Use of GMSE: Model case studies

![](African_Bush_Elephant.jpg)  
**Image:** [M. M. Karim](https://en.wikipedia.org/wiki/User:Muhammad_Mahdi_Karim)

![](barnacle_goose.JPG)  
**Image:** [Raju Kasambe](https://commons.wikimedia.org/wiki/File:Barnacle_Geese_Branta_leucopsis_London_UK_(4).JPG)

## Use of GMSE: Theoretical questions

![](man_freq.png)
